### 2021-11-01 writing/Haedong-SimGuided-Kv-Modeling.Rmd
Writing - Introduction

---
title: Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes
author: "Haedong Kim, Hui Yang, Andrew R. Ednie, and Eric S. Bennet"
date: "10/18/2021"
output: pdf_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Abstract

Electrical signaling initiates and controls the contraction of cardiac muscles. Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes, so they have been important subjects of research in cardiac electrophysiology. Among major cardiac VGICs, K\textsuperscript{+} channel (K\textsubscript{v}) has a distinctive feature in that it has multiple isoforms. Each isoform produces a different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}). Despite the importance of studying each K\textsubscript{v} isoform rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments. Hence, we need a data assimilation method that combines mathematical models with experimental data, I\textsubscript{Ksum}, to estimate individual K\textsuperscript{+} currents. Exponential fitting is a traditional method to estimate the current functions of K\textsubscript{v} isoforms from I\textsubscript{Ksum} data based on the observation that K\textsuperscript{+} currents have the shape of an exponential decay function. It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents. However, this classical method does not consider the kinetics of ion channels at the cellular level but only the shape of each I\textsubscript{Ksum} recording. On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities in the heart. This study proposes a data assimilation method based on simulation of K\textsubscript{v} kinetics that combines in-vitro experimental results I\textsubscript{Ksum} with in-silico models of K\textsubscript{v} isoforms by calibrating the computer models. The proposed method differs from the exponential fitting in a way that it estimates the characteristics of current functions and underlying kinetics. In addition, the suggested approach assimilates I\textsubscript{Ksum} recordings at cellular level rather than analyzing each recording independently. An accompanying graphical user interface (GUI) is developed to provide a user-friendly environment to test the proposed method for practitioners. We conducted a case study with our in-vitro experimental data that investigated the effect of reduced glycosylation on K\textsubscript{v}.

**Keywords:** Cardiac electrophysiology models, data assimilation, model calibration, K\textsubscript{v} isoforms, mouse cardiomyocytes

## Introduction

Electrical signaling plays a vital role in the heart conduction system that controls electrical excitation propagation through the heart, which triggers and synchronizes the contraction of a myriad of cardiac muscle cells [@veeraraghavan2014mechanisms]. In turn, this electrical stimulation is responsible for the heart to pump blood throughout the body properly. Cardiac voltage-gated ion channels (VGICs) are membrane proteins in a cardiomyocyte that control specific ions to flow through the membrane [@bezanilla2005vgic]. They are fundamental functional units of electrical activities in cardiomyocytes as the electrical excitation results from the orchestrated dynamics of different types of ionic currents controlled by the gating kinetics of VGICs as shown in Figure \ref{fig:variety_currents} [@nerbonne2003molecular; @grant2009cardiac]. This electrical simulations is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels. Na\textsubscript{v} are responsible for AP initiation (depolarization phase), Ca\textsubscript{v} are responsible for the prolonged depolarization phase (``plateau'') particularly in larger species like human, while various isoforms of K\textsubscript{v}s are collectively responsible for AP deactivation (repolarization phase).
```{r variety_currents, fig.align='center', fig.lp='fig:', fig.cap='A schematic illustration of action potentials in mouse ventricular cardiomyocytes and underlying ionic currents of Na\textsubscript{v}, Ca\textsubscript{v}, and K\textsubscript{v} isoforms'}
knitr::include_graphics('figs/variety_currents.pdf')
```

In contrast to Na\textsubscript{v} and Ca\textsubscript{v}, K\textsubscript{v} have a distinctive feature that they have multiple isoforms. Each isoform produces different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) and contributes to different portions of the AP repolarization. To be specific, as shown in Figure \ref{fig:variety_currents}, there are various K\textsubscript{v} isoforms and currents they produce: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) [@brouillette2004functional; @liu2011dissection]. Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos}, only one of them is presented and referred to as I\textsubscript{Kto} because I\textsubscript{Ktof} is only found in ventricular apex myocytes, while I\textsubscript{Ktos} is reported to appear only in the septum [@xu1999four]. There are several K\textsubscript{v} isoforms whose exact roles in shaping the cardiac AP are still subject of intensive research. Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments.

Hence, we need a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}. Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations. To be specific, for our case, a data assimilation method is required that couples mathematical models of K\textsuperscript{+} models with experimental data of I\textsubscript{Ksum} using whole-cell patch-clamp recordings to estimate various K\textsuperscript{+} currents. Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation~\ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters. I\textsubscript{kss} is simplified as constant $A\textsubscript{Kss}$. It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents and served as a standard method to analyze the sum of K\textsuperscript{+} currents [@brunet2004heterogeneous; liu2011dissection]. However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level.
\begin{equation}
    \text{I}\textsubscript{Ksum} = \sum_{i} A_{i} e^{-t/\tau_{i}} + A\textsubscript{Kss}, \; i \in I \subseteq \{\text{Kto, } \text{Kslow1, } \text{Kslow2}\} 
\end{equation}

On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities at the cellular level that enable the quantitative understanding of biophysical functions such as gating kinetics of VGICs in the behavior of cardiomyocytes [@winslow2011integrative]. Therefore, this paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level. K\textsubscript{v} models are developed on biophysical principles of VGICs that are encoded in terms of mathematical equations. These equations have parameters that act like ``knobs'', allowing to control the behavior of the models. Tunning computer models to couple with experimental data requires trial-and-error investigations and domain knowledge in both the kinetics of cardiomyocytes and mathematics of computer models. Since modeling of cardiac electrophysiology is an interdisciplinary field that incorporates biology and engineering, calibration of computer models and interpretation of modeling results can be challenging for researchers in biology or medicine communities without engineering training. This hurdle hinders researchers in physiology from utilizing computer models to analyze their experimental data. A graphical user interface (GUI) is provided in which researchers can easily calibrate K\textsubscript{v} models to their data and visualize the modeling results. A new model calibration method will allow researchers to utilize computer models to analyze complex potassium dynamics in cardiomyocytes without engineering background. A GUI will provide an interactive environment to use the calibration method and analyze the results easily with visualizations.

## Material and Methods

## Results

## Conclusions

## References