### 2021-11-17 writing/Haedong-SimGuided-Kv-Modeling.Rmd
Abstract revised

---
title: Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes
author: "Haedong Kim, Hui Yang, Andrew R. Ednie, and Eric S. Bennet"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Abstract

* **Motivation** - Among major cardiac voltage-gated ion channels (VGICs), K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature that they have various isoforms. Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments. Data analysis tools are used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.

* **Gaps & Needs** - A traditional approach is based on the data-driven curve fitting method that assumes the functional forms for the shape of individual K\textsuperscript{+} currents. This method directly fits individual K\textsuperscript{+} currents to experimental I\textsubscript{Ksum} recordings by searching the shape parameters of the current functions that have the best fit. It has been proven that the curve fitting approach describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents. However, the mathematical form of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, the curve fitting method handles experimental I\textsubscript{Ksum} recordings from the same cell independently, so it lacks the cellular-level analysis.

* **Objective** - As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms based on computer models of K\textsuperscript{+} currents that can simulate underlying gating kinetics of the currents.

* **Approach**
  1. Computer models are designed with calibration parameters that allow us to control the behavior of the ion channel models.
  2. A fractional factorial design is utilized to identify the parameters having significant effects on the current shape.
  3. Then, various non-linear optimization methods were applied and compared to calibrate the selected parameters to minimize the discrepancies between the simulated currents and I\textsubscript{Ksum} recordings from the same cell.
  4. Further, we develop a graphical user interface (GUI) to provide a user-friendly environment to test the proposed method for practitioners unfamiliar with coding. 
  5. We conduct a case study with our in-vitro experimental data investigating the effect of reduced glycosylation on K\textsubscript{v}.

* **Experiment & Performance** - The proposed framework outperforms the curve fitting method for minimizing discrepancies between experimental and estimated total K\textsuperscript{+} currents. We also provide the kinetics modeling results that are not possible in the traditional method.

* **Significance** - Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms.

**Keywords:** Cardiac electrophysiology models, data assimilation, model calibration, K\textsubscript{v} isoforms, mouse cardiomyocytes

## Introduction

Electrical signaling plays a vital role in the heart conduction system that controls electrical excitation propagation through the heart, which triggers and synchronizes the contraction of a myriad of cardiac muscle cells [@veeraraghavan2014mechanisms]. In turn, this electrical stimulation is responsible for the heart to pump blood throughout the body properly. Cardiac voltage-gated ion channels (VGICs) are membrane proteins in a cardiomyocyte that control specific ions to flow through the membrane [@bezanilla2005vgic]. They are fundamental functional units of electrical activities in cardiomyocytes as the electrical excitation results from the orchestrated dynamics of different types of ionic currents controlled by the gating kinetics of VGICs as shown in Figure \ref{fig:variety_currents} [@nerbonne2003molecular; @grant2009cardiac]. This electrical simulations is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels. Na\textsubscript{v} are responsible for AP initiation (depolarization phase), Ca\textsubscript{v} are responsible for the prolonged depolarization phase (``plateau'') particularly in larger species like human, while various isoforms of K\textsubscript{v}s are collectively responsible for AP deactivation (repolarization phase).
```{r variety_currents, fig.align='center', fig.lp='fig:', fig.cap='A schematic illustration of action potentials in mouse ventricular cardiomyocytes and underlying ionic currents of Na\textsubscript{v}, Ca\textsubscript{v}, and K\textsubscript{v} isoforms'}
knitr::include_graphics('figs/variety_currents.pdf')
```

In contrast to Na\textsubscript{v} and Ca\textsubscript{v}, K\textsubscript{v} have a distinctive feature that they have multiple isoforms. Each isoform produces different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) and contributes to different portions of the AP repolarization. To be specific, as shown in Figure \ref{fig:variety_currents}, there are various K\textsubscript{v} isoforms and currents they produce: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) [@brouillette2004functional; @liu2011dissection]. Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos}, only one of them is presented and referred to as I\textsubscript{Kto} because I\textsubscript{Ktof} is only found in ventricular apex myocytes, while I\textsubscript{Ktos} is reported to appear only in the septum [@xu1999four]. There are several K\textsubscript{v} isoforms whose exact roles in shaping the cardiac AP are still subject of intensive research. Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments.

Hence, we need a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}. Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations. To be specific, for our case, a data assimilation method is required that couples mathematical models of K\textsuperscript{+} models with experimental data of I\textsubscript{Ksum} using whole-cell patch-clamp recordings to estimate various K\textsuperscript{+} currents. Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation \ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters. I\textsubscript{kss} is simplified as constant $A\textsubscript{Kss}$. It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents and served as a standard method to analyze the sum of K\textsuperscript{+} currents [@brunet2004heterogeneous; liu2011dissection]. However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level.
\begin{equation}
\label{eq:sumexp_model}
    \text{I}\textsubscript{Ksum} = \sum_{i} A_{i} e^{-t/\tau_{i}} + A\textsubscript{Kss}, \; i \in I \subseteq \{\text{Kto, } \text{Kslow1, } \text{Kslow2}\} 
\end{equation}

On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities at the cellular level that enable the quantitative understanding of biophysical functions such as gating kinetics of VGICs in the behavior of cardiomyocytes [@winslow2011integrative]. Therefore, this paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level. K\textsubscript{v} models are developed on biophysical principles of VGICs that are encoded in terms of mathematical equations. These equations have parameters that act like ``knobs'', allowing to control the behavior of the models. Tunning computer models to couple with experimental data requires trial-and-error investigations and domain knowledge in both the kinetics of cardiomyocytes and mathematics of computer models. Since modeling of cardiac electrophysiology is an interdisciplinary field that incorporates biology and engineering, calibration of computer models and interpretation of modeling results can be challenging for researchers in biology or medicine communities without engineering training. This hurdle hinders researchers in physiology from utilizing computer models to analyze their experimental data. A graphical user interface (GUI) is provided in which researchers can easily calibrate K\textsubscript{v} models to their data and visualize the modeling results. A new model calibration method will allow researchers to utilize computer models to analyze complex potassium dynamics in cardiomyocytes without engineering background. A GUI will provide an interactive environment to use the calibration method and analyze the results easily with visualizations.

## Material and Methods

### Computer Models

As shown in Figure \ref{fig:variety_currents}, there are four primary currents yield the most portion of the total K\textsuperscript{+} current: I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}. Computer models of the four currents are developed based on the Hodgkin-Huxley type modeling principle. The Hodgkin-Huxley model treats ion channels as electrical components in a circuit and represents the permeability changes of ion channels as the electrical conductance changes [@fang2021memristive]. In the typical Hodgkin-Huxley modeling, K\textsuperscript{+} current is defined by Equation \ref{eq:hh_current} 
\begin{equation}
\label{eq:hh_current}
  I=g_Ka^{n}i^{m}(V-E_K),
\end{equation}
where $g_K$ is the maximum conductance, $a$ and $i$ are state variables of gating kinetics, $n$ and $m$ are non-negative integers, $V$ is the membrane potential, and the $E_K$ is the Nernst potential.

### Sensitivity Analysis and Calibration Parameters

figures of half-normal probability plots and selected parameters and their principles (it was not purely data-driven but I selected a few subjectively)

### Model Calibration

objective function and optimization methods

### Graphical User Interface

we develope gui

## Results

## Conclusions

## References

<div id="refs"></div>

## Appendix

full equations of potassium models