### 2022-03-23 KvAppUserManual.Rmd
User Guide

---
title: "User Manual for K~v~App"
author: "Haedong Kim"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started

Although K~v~App is developed in MATLAB, it does not require MATLAB installed on the machine. Instead, running `MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without MATLAB. A machine needs to be connected to the Internet to complete the installation.

## Modules and Functions

KvApp

## Example

example with 4.5-sec dat and 25-sec data