### 2022-03-27 KvApp_guide.Rmd
Getting started

---
title: "User Guide for KvApp"
author: "Haedong Kim"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started

There are four associated files as shown in Figure 1 below: `KvAppBeta`, `MyAppInstaller_mcr`, `MyAppInstaller_web`, and `splash`.

![Consisting files of the applications.](KvApp_guide_img/folder_struct.png){width=50%}

`KvAppBeta` is the file that runs the application, but you will get the error message in Figure 2 if the correct version of MATLAB is not installed on the machine.

![Error message when the right version of MATLAB is not installed on the target machine.](KvApp_guide_img/runtime_error.png){width=50%}

`MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without having MATLAB installed. Your computer needs to be connected to the Internet to complete the installation. If the Internet connection is not available, `MyAppInstaller_mcr` will install the necessary programs locally. A few screenshots below show the installation process.

![](KvApp_guide_img/install1.png){width=50%} ![](KvApp_guide_img/install4.png){width=50%}
![](KvApp_guide_img/install5.png){width=50%} ![](KvApp_guide_img/install6.png){width=50%}

The other file `splash` is just an image file for the application icon.

## Modules and Functions

KvApp

## Example

example with 4.5-sec dat and 25-sec data