### 2022-11-02 writing/draft.tex
Draft revision

\documentclass[11pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

%%%%%%% IISE Transactions margin specifications %%%%%%%%%%%%%%%%%%%
% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\normalfont\fontfamily{phv}\fontsize{16}{19}\bfseries}}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\fontfamily{phv}\fontsize{14}{17}\bfseries}}
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                                    {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\normalsize\fontfamily{phv}\fontsize{14}{17}\selectfont}}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% IISE Transactions package list %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib} %comment out if you do not have the package
\usepackage{url} % not crucial - just used below for the URL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% Author package list and commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Here are some examples %%%%%%%%%%%%%%
%	\usepackage{amsfonts, amsthm, latexsym, amssymb}
%	\usepackage{lineno}
%	\newcommand{\mb}{\mathbf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\def\spacingset#1{\renewcommand{\baselinestretch}%
		{#1}\small\normalsize} \spacingset{1}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\if0\blind
	{
		\title{\bf Model-guided Concurrent Data Assimilation for Calibrating Cardiac Ion-channel Kinetics}
		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
		$^a$ Department of Industrial and Manufacturing Engineering, \\
		The Pennsylvania State University, University Park, PA USA \\
        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
        Wright State University, Dayton, OH USA}
		\date{}
		\maketitle
	} \fi

	\if1\blind
	{
        \title{\bf \emph{IISE Transactions} \LaTeX \ Template}
		\author{Author information is purposely removed for double-blind review}

\bigskip
		\bigskip
		\bigs   kip
		\begin{center}
			{\LARGE\bf \emph{IISE Transactions} \LaTeX \ Template}
		\end{center}
		\medskip
	} \fi
	\bigskip

\begin{abstract}
Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes. Different K\textsubscript{v} isoforms and their currents contribute to various phases of repolarization, e.g., the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1). Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose and estimate different currents of K\textsubscript{v} isoforms. Most existing studies separate and make inference of K\textsubscript{v} kinetics via a curve-fitting procedure (e.g., tri-exponential fitting), but are limited in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be decomposed and fitted for one clamp voltage at a time. In this paper, we develop a novel \textit{concurrent data assimilation} framework that calibrates biophysics-based computer models to delineate kinetic dynamics of various K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously. The proposed method is evaluated and validated with voltage-clamp I\textsubscript{Ksum} recordings from our \textit{in-vitro} experiments on both wild-type and glyco-altered myocytes. Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes disease-altered perturbations to various K\textsubscript{v} isoforms. In addition, we develop a graphical-user-interface (GUI) software that provides an enabling tool to medical scientists for model calibration with voltage-clamp data. The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in variety of heart diseases.
\end{abstract}

\noindent%
{\it Keywords:} Data assimilation, simulation model, ion channels, congenital glycosylation disorder, model calibration

%\newpage
\spacingset{1.5} % DON'T change the spacing!

\section{Introduction}
Cardiac voltage-gated potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart. K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time representing the net electrical activity in a cardiomyocyte \citep{ravens2008role}. The shape and duration of the AP are primarily determined by various K\textsubscript{v} isoforms, contributing to different phases of repolarization with unique biophysical properties as illustrated in \textbf{Figure~\ref{fig:kv_ap}} \citep{nerbonne2005molecular}. Diverse distribution is a unique feature of K\textsubscript{v}. Dominant isoforms exist for other major voltage-gated ion channels (VGICs) such as Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}. Repolarization begins with the rapid inactivation of Na\textsubscript{v}, reducing inward Na\textsuperscript{+} flux, and fast open and close of K\textsubscript{v} isoforms, generating rapidly inactivating K\textsuperscript{+} current. There are other K\textsubscript{v} that remain open producing slowly inactivating currents, but in large species such as human, this phase 2 is characterized by prolonged repolarization due to intracellular Ca\textsuperscript{2+} flux (``plateau''). During phase 3, Ca\textsubscript{v} close and the cell begins to repolarize rapidly (``rapid repolarization''). In small species such as mice, phase 2 and 3 are not very distinctive. Hence, in the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating delayed rectifier-type currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively \citep{brunet2004heterogeneous}. Aberrant repolarization can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{montpetit2009regulated}. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kv_ap.pdf}
    \caption{Ventricular action potential in mouse cardiomyocytes and the diverse K\textsubscript{v} isoforms and other ion channels contributing to shaping the AP.}
    \label{fig:kv_ap}
\end{figure}

However, only the collective activities of K\textsubscript{v} isoforms can be measured in the form of currents using whole-cell voltage-clamp recording (I\textsubscript{Ksum}) due to their slightly overlapping voltage dependence of gating \citep{brouillette2004functional}. Hence, there is a need to decompose I\textsubscript{Ksum} into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological mechanisms in disease-altered cardiomyocytes. Data assimilation is a method of data analysis that combines experimental data with mathematical/computational models to make inferences on information of interest, such as individual K\textsuperscript{+} currents. Most existing studies separate individual K\textsuperscript{+} currents mathematically via a curve-fitting procedure, assuming a curve of I\textsubscript{Ksum} as the summation of exponential decay functions and a constant term; each represents the shape of individual K\textsuperscript{+} currents \citep{brunet2004heterogeneous}. However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition is sorely determined by the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) although a set of I\textsubscript{Ksum} traces are recorded with varying clamp voltages for a cell to investigate the voltage dependence of gating kinetics, only a single I\textsubscript{Ksum} trace is decomposed and fitted to individual K\textsuperscript{+} currents for one clamp voltage. ion methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{Ksum} traces for a better understanding of the ion channels to provide insights into mechanism of cardiomyocytes and clinical applications, such as drug design.

Therefore, this paper proposes a new approach of \textit{concurrent data assimilation} for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{Ksum} traces measured in the same cell. First, \textit{in-silico} models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents. It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents. \textbf{Figure~\ref{fig:exp-vs-com}} shows two different approaches to assimilate the experimental data to extract information of K\textsubscript{v} isoforms. Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant impacts on the current generation. Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in various perspectives. Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings concurrently. Last, low-dimensional embedding is utilized to visually inspect the differences of the calibrated parameters in a collective way across the cells. We apply the proposed framework to glyco-altered I\textsubscript{Ksum} recordings from our previous studies. Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}. Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}. To be specific, reduced N-glycosylation in mouse cardiomyocytes is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/exp-vs-com.pdf}
    \caption{Comparison between the traditional curve-fitting approach and the proposed model-guided data assimilation of \textit{in-vitro} experimental data of I\textsubscript{Ksum} recordings.}
    \label{fig:exp-vs-com}
\end{figure}

Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} isoform  currents and underlying kinetics of both the wild-type and disease-altered groups and successfully captures cellular variability. \textit{In-silico} modeling provides rich information that \textit{in-vitro} data cannot alone. For example, in our previous studies and this paper, computer models are used to verify experimental results, estimate channel kinetics, and show how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}. \textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases. Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to medical scientists for investigating K\textsubscript{v} channelopathy, which is available as online supplemental material. The main advantage of the proposed approach is that it analyzes multiple I\textsubscript{Ksum} recordings concurrently based on the biophysical computer models, providing strong potential for modeling the kinetics and gating mechanism of the cardiac ion channels for studying heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{Ksum} recordings based on biophysical principles rather than curve-fitting procedures.
    \item We developed a concurrent data assimilation routine that harnesses a set of I\textsubscript{Ksum} traces simultaneously rather than a single current trace independently one-by-one, resulting in comprehensive cellular-level modeling.
    \item The proposed approach generates separate K\textsubscript{v} models and visualize the distributions of the kinetic parameters for each cell, which allows analyzing the cellular variability. 
    \item Our method captures differences in channel kinetics of the wild-type and the disease-altered groups.
\end{itemize}
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/overall_diagram.pdf}
    \caption{Conceptual diagram to illustrate a data assimilation framework.}
    \label{fig:framework_diagram}
\end{figure}

\section{Research Background}
\subsection{Data Assimilation and Calibration of Cardiac Models}
The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself. Cardiac electrophysiology has contributed significantly to our understanding of heart and disease-related modifications, such as congenital disorders of glycosylation (CDG). Advances in laboratory techniques allow studying molecular-level activities in cardiomyocytes via measuring currents of ion channels, which are functional units of the electrical conduction system of the heart. For example, whole-cell current recordings are used to show the roles of a certain gene, encoding $\alpha$ subunits that form Na\textsubscript{v} in human cardiomyocytes, which causes a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}. Analysis of ionic currents showed pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities \citep{ednie2013sialicNav1, ednie2015sialicKv}; O-glycosylation modulations cause aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}. However, \textit{in-vitro} experiments alone are limited to investigate channel activities. There is a need to assimilate experimental data to estimate information that cannot be observed directly. Traditionally, a curve-fitting method is used to make inferences, which assumes a functional form of information to estimate (e.g., individual currents of K\textsubscript{v} isoforms) \citep{liu2011dissection}.

However, the curve-fitting approach has limitations in its ability to assimilate the experimental data, because it only considers the shape of information to infer rather than reflecting underlying kinetics and interactions of various gating processes. Further, as curve fitting is only based on the shape of a single curve to fit, the traditional approach is not able to analyze a set of current recordings that are supposed to be analyzed together. In contrast, computer models of the heart consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations. Mathematical and computational implementations of the heart allow studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}. For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}. Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}. In the computational modeling literature, discrete-event simulation has been a primary focus, which only models events that occur apart over time. However, computer models of ion channels simulate gating kinetics that changes continuously, expressed mathematically in differential equations. The complex nonlinear structure of cardiac computer models also challenges calibration. However, little work has been done to calibrate continuous simulation-guided model.

\subsection{Congenital Disorders of Glycosylation}
Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}. It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}. However, the underlying pathological mechanism still remains elusive.

We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology. Electrical signaling is orchestrated activities of a variety of ion channels and transporters. Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}. Glycosylation is a multi-step process and usually ends with sialic acid added. We also reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}. Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.

However, most of existing works utilized the curve-fitting method, e.g., bi-exponential or tri-exponential fitting, to assimilate the \textit{in-vitro} data from glycosylayion experiments. Curve-fitting provides estimates of characteristic curves (e.g., steady-state activation or inactivation) and statistics (e.g., time constants) for data-driven insights into CDG diseases. Few, if any, previous approaches have leveraged computer models to calibrate and delineate kinetic dynamics of various K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.

\section{Research Methodology}
\subsection{Computer Models of Potassium Channel Isoforms}
Mouse models have been very useful methods for studying human cardiac electric signaling \citep{nerbonne2004studying,milani2014small,kim2022irx5}. Computer models of mouse cardiomyocytes simulate important biophysical properties of ion channels and molecular dynamics in producing the AP. In this paper, the AP is modeled as a differential equation of transmembrane currents and stimulus current I\textsubscript{stim} as below in Equation~\ref{eq:ap} where $C_{m}$ is the membrane capacitance, $t$ is time.
\begin{equation}
    \label{eq:ap}
    \begin{split}
    -C_{m}\frac{dV}{dt} &= \mathrm{I}_{\mathrm{CaL}}+\mathrm{I}_{\mathrm{p}(Ca)}+\mathrm{I}_{\mathrm{NaCa}}+\mathrm{I}_{\mathrm{Cab}}+\mathrm{I}_{\mathrm{Na}}+\mathrm{I}_{\mathrm{Nab}}+\mathrm{I}_{\mathrm{NaK}}+\mathrm{I}_{\mathrm{Cl},Ca} \\
    &+\mathrm{I}_{\mathrm{Kto}}+\mathrm{I}_{\mathrm{Kslow1}}+\mathrm{I}_{\mathrm{Kslow2}}+\mathrm{I}_{\mathrm{Kss}}+\mathrm{I}_{\mathrm{Ks}}+\mathrm{I}_{\mathrm{Kr}}+\mathrm{I}_{\mathrm{K1}}+\mathrm{I}_{\mathrm{stim}}
    \end{split}
\end{equation}
There are 15 transmembrane currents in this AP model: the L-type calcium current (I\textsubscript{CaL}), the calcium pump current (I\textsubscript{p(Ca)}), the Na\textsuperscript{+}/Ca\textsuperscript{2+} exchange current (I\textsubscript{NaCa}), the calcium background current (I\textsubscript{Cab}), the fast Na\textsuperscript{+} current (I\textsubscript{Na}), the background Na\textsuperscript{+} current (I\textsubscript{Nab}), the Na\textsuperscript{+}/K\textsuperscript{+} pump current (I\textsubscript{NaK}), the Ca\textsuperscript{2+}-activated Cl\textsuperscript{-} current (I\textsubscript{Cl,Ca}), the rapidly inactivating transient outward K\textsuperscript{+} current (I\textsubscript{Kto}), the two ultra-rapidly activating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}), the slow delayed rectifier K\textsuperscript{+} current (I\textsubscript{Ks}), the rapid delayed rectifier K\textsuperscript{+} current (I\textsubscript{Kr}), and the time independent K\textsuperscript{+} current (I\textsubscript{K1}). Detailed implementation of mouse AP and VGIC models can be found in \cite{bondarenko2014compartmentalized,asfaw2020compartmentalized}.

The three types of K\textsuperscript{+} currents have larger magnitude compared to the other K\textsuperscript{+} currents: the rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), the delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}). The rapidly inactivating transient outward currents have two components I\textsubscript{Kto,f} and I\textsubscript{Kto,s}, but the later is essentially absent in apex cardiomyocytes. We only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because we focus on apex myocytes in this investigation. However, the transient outward current can be easily modified according to the region of ventricular cardiomyocytes. Therefore, we model K\textsubscript{v} isoforms that conduct these dominant currents: K\textsubscript{v}4.2 (I\textsubscript{Kto}), K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}), and K\textsubscript{2P} family (I\textsubscript{Kss}). \textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{Ksum} traces (I\textsubscript{Ksum}) given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$. \textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of voltage-clamp protocol and consequent change in I\textsubscript{Ksum}. Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach. In this approach, K\textsuperscript{+} current is defined by Equation~\ref{eq:expl_fitting} with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$, and a set of them is estimated that have the best fit with experimental I\textsubscript{Ksum} data.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/exemplary_kv_currents.pdf}
    \caption{Example of voltage-clamp K\textsuperscript{+} currents. (A) Voltage step of 0 mV from holding potential -70 mV. (B) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}. (C) Range of voltage-clamp protocol from -30 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
    \label{fig:kcurrent_example}
\end{figure}
\begin{equation}
    \label{eq:expl_fitting}
    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}
\end{equation}

Unlike the simple current functions, computer models of K\textsubscript{v} isoforms can simulate detailed gating kinetics. We developed them with Hodgkin-Huxley modeling, which has two gating variables controlling conductance, a popular modeling scheme in various species \citep{ten2004model,mahajan2008rabbit}. For example, K\textsubscript{v} isoforms that conducts current I\textsubscript{K} can be defined by Equation~\ref{eq:hh_ex}
\begin{equation}
    \label{eq:hh_ex}
    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}})
\end{equation}
where $G_{\mathrm{K}}$ is the maximum conductance, $a^{n}$ and $i^{m}$ are the gating variables for $n,m \in \mathbb{N}$, $V$ is the transmembrane potential, and $E_{\mathrm{K}}$ is the K\textsuperscript{+} Nernst potential. $V-E_{\mathrm{K}}$ is the driving force of the ion movement. Important components in this equation are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$. These processes are governed by first-order kinetics and its voltage-dependent transition rates $\alpha$ and $\beta$. $\alpha$ is the rate at which a gate in closed state opens, whereas $\beta$ is the rates at which a gate in open state closes. Equation~\ref{eq:gv} shows a schematic relationship of this gating kinetics.
\begin{align}
    \label{eq:gv}
    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i
\end{align}

The two biophysical processes can be modeled using differential equations in two ways as follows:
\begin{align}
    \frac{da}{dt} &=\alpha_{a}(1-a)-\beta_{a}a   &\frac{di}{dt} &=\alpha_{i}(1-i)-\beta_{i}i \\
    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}}
\end{align}
where $a_{\infty}$ and $i_{\infty}$ are the steady-state values to which $a$ and $b$ converge; $\tau_{a}$ and $\tau_{i}$ are time constants determine the convergence speed defined by
\begin{align}
    a_{\infty} &= \frac{\alpha_{a}}{\alpha_{a}+\beta_{a}} & i_{\infty} &=  \frac{\alpha_{i}}{\alpha_{i}+\beta_{i}} \\
    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}
\end{align}
The steady-state values and time constants can be defined directly by functions of voltage without transition rates in some cases. These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters that control the behavior of the kinetics of an ion channel.

\subsubsection{In-silico Modeling of K\textsubscript{v}4.2 (I\textsubscript{Kto})}
The rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, is characterized by a sharp upstroke during activation and subsequent rapid inactivation. It mainly contributes to the peak at the very beginning of activation in I\textsubscript{Ksum}. I\textsubscript{Kto} is defined by
\begin{align}
    &\mathrm{I}_{\mathrm{Kto}} = G_{\mathrm{Kto}}a_{\mathrm{Kto}}^{3}i_{\mathrm{Kto}}(V-E_{\mathrm{K}}) \\
    &\frac{da_{\mathrm{Kto}}}{dt} = \alpha_{a}(1-a_{\mathrm{Kto}}) - \beta_{a}a_{\mathrm{Kto}} \\
    &\frac{di_{\mathrm{Kto}}}{dt} = \alpha_{i}(1-i_{\mathrm{Kto}}) - \beta_{i}i_{\mathrm{Kto}} \\
    &\alpha_{a} = p_{7}e^{p_{5}(V+p_{1})} \label{eq:ikto_alpha1} \\
    &\beta_{a}= p_{8}e^{-p_6(V+p_{1})} \\
    &\alpha_{i} = \frac{p_{9}e^{-(V+p_{2})/p_{4}}}{1+p_{10}e^{-(V+p_{2}+p_{3})/p_{4}}} \\
    & \beta_{i} = \frac{p_{11}e^{(V+p_{2}+p{3})/p_{4}}}{1+p_{12}e^{(V+p_{2}+p_{3})/p_{4}}} \label{eq:ikto_beta2}
\end{align}
There are two gating variables $a_{\mathrm{Kto}}$ and $i_{\mathrm{Kto}}$ responsible for activation and inactivation, respectively. Their kinetics are governed by transition-rate functions from Equation~\ref{eq:ikto_alpha1} to Equation~\ref{eq:ikto_beta2}. Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.

\subsubsection{In-silico Modeling of K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}) and K\textsubscript{2P} (I\textsubscript{Kss})} \label{s:method.model2}
There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are very rapidly activating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively. As shown in \textbf{Figure~\ref{fig:kcurrent_example}}, both rectifier currents inactivate slower and have smaller magnitude than I\textsubscript{Kto}, but I\textsubscript{Kslow2} decays more gradually than I\textsubscript{Kslow1}. The non-inactivating steady-state current, which conducted through K\textsubscript{2P} family, remains constant during the voltage-clamp recording. These three currents collectively address the most part of the decaying phase of I\textsubscript{Ksum}. We assume that I\textsubscript{Kslow1} and I\textsubscript{Kslow2} have the same activation gating variable, and I\textsubscript{Kss} has similar activation behavior with a slightly different rate to keep the model as simple as possible to reduce the structural risk of overfitting.

I\textsubscript{Kslow1} is modeled without transition-rate functions as opposed to I\textsubscript{Kto}. Its gating variables, activation $a_{\mathrm{Kslow1}}$, and inactivation $i_{\mathrm{Kslow1}}$ are defined directly by steady-state ($a_{ss}$ and $i_{ss}$) and time-constant functions ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) as follows.
\begin{align}
    &\mathrm{I}_{\mathrm{Kslow1}} = G_{\mathrm{Kslow1}}a_{\mathrm{Kslow1}}i_{\mathrm{Kslow1}}(V-E_{\mathrm{K}}) \\
    &\frac{da_{\mathrm{Kslow1}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kslow1}}}{\tau_{a}^{(1)}} \\
    &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(1)}} \\
    &a_{ss} = \frac{1}{1+e^{-(V+p_{1})/p_{4}}} \label{eq:ass} \\
    &i_{ss} = \frac{1}{1+e^{(V+p_{2})/p_{5}}} \label{eq:iss} \\
    &\tau_{a}^{(2)} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
    &\tau_{i}^{(2)} = p_{10} - p_{8}i_{ss}
\end{align}

I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1}, and the time-constant function of the inactivation $i_{\mathrm{Kslow2}}$ that contains the same steady-state function $i_{ss}$ in I\textsubscript{Kslow1}. As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows. 
\begin{align}
    &\mathrm{I}_{\mathrm{Kslow2}} = G_{\mathrm{Kslow2}}a_{\mathrm{Kslow2}}i_{\mathrm{Kslow2}}(V-E_{\mathrm{K}}) \\
    &a_{\mathrm{Kslow2}} = a_{\mathrm{Kslow1}} \\
    &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(2)}} \\
    &\tau_{i}^{(3)} = p_{2} - p_{1}i_{ss}
\end{align}

I\textsubscript{Kss} does not have an inactivation variable as it is non-inactivating current. It shares the same steady-state function for activation $a_{ss}$ with the rectifier currents but have a separate time-constant function to address the different activation rate. I\textsubscript{Kss} is modeled as
\begin{align}
    &\mathrm{I}_{\mathrm{Kss}} = G_{\mathrm{Kss}}a_{\mathrm{Kss}}(V-E_{\mathrm{K}}) \\
    &\frac{da_{\mathrm{Kss}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kss}}}{\tau_{a}^{(3)}} \\
    &\tau_{a}^{(4)}= \frac{p_{2}}{e^{p_{1}(V+p_{3}^\prime)}+e^{-p_{1}(V+p_{3}^\prime)}} + p_{3}
\end{align}
Note that $p_{3}^\prime$ is equal to $p_{3}$ in I\textsubscript{Kslow1}.

\subsection{Concurrent Assimilation of Functional Data}
Data assimilation is a systematic procedure to find the optimal configuration and state of computational/mathematical models by coupling it with experimental data. Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation. The output of physical experiments $y^{\mathcal{D}}(x)$, given input $x$, inevitably contains errors for various reasons, such as noise in measurement or experimental environment. Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in Equation~\ref{eq:rd_relation}, where $\epsilon$ is the error term. 
\begin{equation}
    \label{eq:rd_relation}
    y^{\mathcal{R}}(x) = y^{\mathcal{D}}(x) + \epsilon
\end{equation}
Let $y^{\mathcal{M}}(x|\theta)$ denote the output from a computer model $\mathcal{M}$, given parameters $\theta$. Assume that there are discrepancies $\delta(x|\theta)$ for the current states of parameters as follows in Equation~\ref{eq:dm_relation}. 
\begin{align}
    \label{eq:dm_relation}
    y^{\mathcal{D}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) \text{, so} \\
    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon
\end{align}
Our goal in data assimilation is to calibrate $\theta$ to find the best model states that minimize $\delta(x|\theta)$, while satisfying biophysical constraints. By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{M}$, which provides two complementary angles to study the real process $\mathcal{R}$.

Traditionally, bi-/tri-exponential functions as described in Equation~\ref{eq:expl_fitting} have served as $y^{\mathcal{M}}(x|\theta)$, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$. Note that the model parameters $\mathrm{A}_{i}(x)$ (amplitude) and $\tau_{i}(x)$ (time constant) depend on input data $x$, which are clamp voltages for whole-cell I\textsubscript{Ksum} recordings. In general, multiple input voltage steps are applied to a myocyte producing a set of I\textsubscript{Ksum} recordings to study the voltage-dependent characteristics of VGICs. However, these exponential functions cannot analyze simultaneously, leading to potential information loss in the cellular-level dynamics of ion channels. In contrast, computer models of VGICs can generate currents given one set of parameters because they are designed by biophysical principles encoding voltage dependence. In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in Equation~\ref{eq:rmse}. RMSE evaluates the goodness-of-fit between a I\textsubscript{Ksum} recording recording and model prediction from the end of a holding potential $t_v^{(h)}$ to the end of voltage step $t_v^{(e)}$. RMSEs from the same cell are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltage $v=1,2,\cdots,n$.
\begin{equation}
    \label{eq:rmse}
    \delta = \sum_{v=1}^{n} \sqrt{\int_{t_v^{(h)}}^{t_v^{(e)}}\frac{(y_v^{\mathcal{D}}(t) - y_v^{\mathcal{M}}(t|\theta))^2}{t_v^{(e)}-t_v^{(h)}}dt}
\end{equation}

This approach differs from previous studies in two-fold. First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method uses computer models of K\textsubscript{v} and incorporate a set of I\textsubscript{Ksum} traces simultaneously for comprehensive cellular-level modeling. Cardiac models enable this concurrent data assimilation, because they generate multiple current traces with different input voltages by simulating underlying gating kinetics. Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}. The proposed model-guided concurrent data assimilation method is summarized and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}}. Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}. We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$. Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube. In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling. Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting. Global optimum is not guaranteed due to the complexity of our models. We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible. Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden. This work is implemented in MATLAB R2022a, and the code is available as online supplemental material.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/flow_chart.pdf}
    \caption{Schematic diagram of (A) the current curve-fitting approach that searches for the best fit of K\textsuperscript{+}-currents curves to I\textsubscript{Ksum} recordings one-by-one and estimates feature kinetics using additional curve fitting if available; (B) the proposed concurrent model-guided data assimilation that calibrates model parameters to multiple I\textsubscript{Ksum} traces and simulates K\textsubscript{+} currents and underlying kinetics from the model.}
    \label{fig:flow_chart}
\end{figure}

\subsection{Model Regularization and Sensitivity Analysis}
The principle of parsimony is critical in model calibration not only for enhancing fitting accuracy and preventing overfitting, but also for improving the interpretability of $\theta$. Excessive flexibility of a model has a risk of causing overfitting that occurs when the model fits data too closely, even including noise and random effects in the data. Besides, as the number of parameters increases, it becomes complicated to interpret the calibration results of parameters. It is worth noting that the models presented in Section 3.1 are designed with structural regularization, in which some parameters and equations are recycled in multiple places to simply the model structure. Specifically, the same parameters that control voltage thresholds, such as $p_{1}$, $p_{2}$, $p_{3}$, and $p_{4}$, are used in equations having a similar function. The steady-state functions (Equation~\ref{eq:ass} and \ref{eq:iss}) are reused in the three rectifier-type currents.

We also performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters. In order to do that, factorial designs are developed in which parameters vary at two levels contrasting their effect on the model output. As illustrated in \textbf{Figure~\ref{fig:current_markers}}, six markers were defined that capture characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments. Each marker represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak. Marker f will equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:current_markers}C} and \textbf{~\ref{fig:current_markers}D}. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/current_markers.pdf}
    \caption{Illustration of the six markers of voltage-clamp K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}. All currents are simulated for illustration, and the labels refer to a) the current magnitude 10 ms after voltage is applied, b) 25\% of the total recording time has elapsed, c) 50\%, d) 75\%, e) the peak magnitude, and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.}
    \label{fig:current_markers}
\end{figure}

We adopted a fractional factorial design of 1024 runs for I\textsubscript{Kto}, which results in resolution VIII. The resolution ensures that the main effects and 2-/3-factor interactions are strongly clear. Full factorial designs are used for the other three currents. The marker points are evaluated at each design point and factorial effects are calculated using the linear model in Equation~\ref{eq:fact_effect} for each current.
\begin{equation}
    \delta^{k} = \beta_{o}^{k} + \sum\limits_{i}\beta_{i}^{k}p_{i}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\},
    \label{eq:fact_effect}
\end{equation}
The least squares method is used to estimate $\beta_{i}^{k}$. Then, half-normal plots are drawn to test significance of the estimated effects.

\subsection{Variability Visualization and Low-dimensional Embedding}
Most cardiac electrophysiology studies utilize statistical tests for the mean to support hypotheses; in turn, the majority of ion channel models are based on population-averaged data. However, fine-grained analyses that consider the variability of cells can provide new insights into data. For example, it is possible there are differences between control and disease-altered groups but also within the groups according to characteristics of cells such as size or age. Hence, there is an urgent need to develop a tool to investigate cellular variability. Because the proposed approach calibrates models for the individual dataset from each cell, it allows us to quantify and visualize cellular variability in the tuned parameters. Histograms are used in this study to visualize distributions of calibration parameters and the effects of disease-related perturbations on the parameter distributions. Although a histogram is a simple yet powerful tool, one of the limitations of a histogram is that it cannot visualize multiple variables more than three on the same plot. It is desirable to inspect how distributional differences in calibration parameters collectively impact inter/intra-cell variability in healthy control and disease groups. Low-dimensional embedding methods transform high-dimensional data into a plane or 3D space while preserving relative locations of data points as much as possible so that the structure of data can be visualized. \textbf{Figure~\ref{fig:embd}} illustrates how a low-dimensional embedding method is used to map the high-dimensional parameter space of data points from wild-type and disease groups into 3D space.  
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/embd.pdf}
    \caption{Diagram illustrating the procedure of low-dimensional embedding to reveal cellular-level structure of the parameter space.}
    \label{fig:embd}
\end{figure}

For this purpose, t-distributed stochastic neighbor embedding (t-SNE) is used \cite{van2008visualizing}. It has proven to be an effective method for visualizing high-dimensional data. t-SNE is a statistical method that constructs two sets of probability distributions $p_{ij}$ and $q_{ij}$ over pairs of data points $i$ and $j$ in a high-/low-dimensional space, respectively. These are probabilities of similarities of data points such that neighboring points have a higher probability while dissimilar points have a lower probability. We first define the conditional probability of $j$ given $i$:
\begin{equation}
    p_{j|i}=
    \begin{cases}\frac{\mathrm{exp}(-d(x_{i},x_{j})^{2}/(2\sigma_{i}^{2}))}{\sum\limits_{k \neq i}\mathrm{exp}(-d(x_{i},x_{k})^{2}/(2\sigma_{i}^{2}))}, \ j \neq i, \\
    0, \ j=i,
    \end{cases}
\end{equation}
where $d(\cdot,\cdot)$ is a distance function such as Euclidean distance, and $\sum\limits_{j} p_{j|i}=1$ for all $i$. Then $p_{ij}$ can be defined by the symmetric property of the joint probabilities:
\begin{equation}
    p_{ij} = \frac{p_{j|i} + p_{i|j}}{2N},
\end{equation}
where $N$ is the number of total data points, and $\sum\limits_{i,j}p_{ij}=1$. It can be calculated from data once the standard deviation $\sigma_{i}$ is given. $\sigma_{i}$ is set in a way that the perplexity of the conditional probability distribution over other data points given $x_{i}$ equals a prefixed value that is a hyperparameter of t-SNE. Let $P_{i}$ denote the conditional probability distribution, then the perplexity of the distribution is
\begin{equation}
    \mathrm{perplexity}(P_{i}) = 2^{H(P_{i})}
\end{equation}
where $H(P_{i})$ is the Shannon entropy of $P_{i}$ defined by
\begin{equation}
    H(P_{i}) = -\sum\limits_{j}p_{j|i}\log_{2}(p_{j|i}).
\end{equation}

Then the probability distribution $q_{ij}$ is defined by the similarity of data points $y_i$ and $y_j$ in a low-dimensional space:
\begin{equation}
    q_{ij} = 
    \begin{cases}\frac{(1+||y_{i}-y_{j}||^{2})^{-1}}{\sum\limits_{k}\sum\limits_{l \neq k}(1+||y_{k}-y_{l}||^{2})^{-1}}, \ j \neq i, \\
    0, \ j=i,
    \end{cases}
\end{equation}
where $\sum\limits_{i,j}q_{ij}=1$. Note that $q_{ij}$ is modeled by a heavy-tailed Student's t-distribution with one degree of freedom, from which the name ``t-SNE'' originates. The objective of t-SNE is to learn $y$ that minimizes discrepancies between $P$ and $Q$, so that the low-dimensional distribution preserves the structure of $p_{ij}$ constructed from the original high-dimensional data. To learn $y$, t-SNE maps $y$ by minimizing the Kullback-Leibler (KL) divergence, which measures the similarity between two probability distributions:
\begin{equation}
    \mathrm{KL}(P||Q) = \sum\limits_{j}\sum\limits_{i \neq j}p{ij}\log \frac{p_{ij}}{q_{ij}}.
\end{equation}
Typically, a gradient descent method is used for the minimization of $\mathrm{KL}(P||Q)$ with respect to $y$.

\section{Experimental Design}
More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}. MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death. To further investigate the role of altered glycosylation in pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}. To be specific, whole-cell I\textsubscript{Ksum} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV. There were 32 observations for each WT and MGAT1KO.

We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation. It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes. A curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in Equation~\ref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents. All three component currents were reduced, but the rectifier current was significantly reduced and slowed notably. Although the \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with altered glycosylation, it was difficult to determine channel kinetics rigorously. For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model. Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{Ksum} recordings. This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics concurrently.

\section{Results}
\subsection{Parameter Screening}
\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six markers. The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points. The farther parameters fall above the straight lines, the more significant impacts they have on the marker. We picked the parameters to be calibrated as a union of the sets of parameters falling above the straight lines on the six half-normal plots. For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$. 
\begin{figure}
    \centering
    \includegraphics{figs/half-norm.pdf}
    \caption{Half-normal plots of factorial effects of parameters on the six maker points.}
    \label{fig:half-norm}
\end{figure}
\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics. We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters. Note that the voltage-dependence parameters in red and green appear multiple times across different equations. This parsimonious model design is intended to maximize the structural regularization to minimize overfitting. In general, voltage-threshold and time-constant-shifting parameters impact the current traces more than others. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/calib_param.pdf}
    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
    \label{fig:calib_param}
\end{figure}

\subsection{Model Fitness}
We used the BFGS algorithm, one of the most popular nonlinear optimization methods, in the proposed model calibration routine because it supports box constraints. A separate accuracy metric is used to measure the goodness-of-fit and potential prediction power of the calibration results rather than utilizing the final objective function value of the calibration routine to report the fitness. The objective function, the sum of RMSEs over voltage steps, does not provide a straightforward interpretation that makes it difficult to measure the performance of the models. $R^{2}$, defined in Equation~\ref{eq:r2}, is the most common prediction accuracy metric ranging from 0 to 1 for linear regression models that represents the proportion of variation in data that is explained by the model. However, this interpretation is not applicable, and its value can be negative for nonlinear models because the variance decomposition in Equation~\ref{eq:var_decomp} does not hold anymore.
\begin{align}
    R^{2} &= 1 - \frac{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2}{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2} \label{eq:r2} \\
    \sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2 &= \sum_{k=1}^{n}(y^{\mathcal{M}}(k|\theta))-\bar{y})^2+\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2 \label{eq:var_decomp}
\end{align}

Therefore, a pseudo-R-squared measure, proposed in \citet{li2019prediction} was used to evaluate the model fitness in this study. This nonlinear $R^{2}$ measure is based on the concept of linear correction of prediction function, allowing the variance decomposition for nonlinear functions, which guarantees straightforward interpretation and normalized scores between 0 and 1 as in the classical $R^{2}$. \textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} present averages of $R^{2}$ values over voltage steps from -30 to 50 mV by 10 mV step size for WT and MGAT1KO, respectively. In most cases, more than 80 \% of variances on average in multiple current recordings are explained by models (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$). The model fitness measures in MGAT1KO are less than WT in general. It is expected that the small current magnitude of MGAT1KO challenges the calibration process. We compared the actual model predictions and experimental I\textsubscript{Ksum} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further. As illustrated in the comparison plot of MGAT1KO at -30 mV clamp voltage, small and noisy current trace results in decreased $R^{2}$, 0.8335. 
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/model_fitness_4half.pdf}
    \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (A) WT and (C) MGAT1KO. Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{Ksum} recordings (in pink) at -30 mV and 50 mV voltage steps of (B) cell 6 in WT and (D) 12 in MGAT1KO.}
    \label{fig:model_fitness}
\end{figure}

\subsection{Kinetics Modeling} \label{s:results.kinetics}
The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior. \textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2. There is a very slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}). Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}. $\tau_{a}^{(1)}$ for MGAT1KO is greater than WT in range of $>$ -20 mV. $\tau_{i}^{(1)}$ for MGAT1KO shifted toward left. The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increased time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kinetics_ikto.pdf}
    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kto} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, and (D) inactivation time constant.}
    \label{fig:kinetics_ikto}
\end{figure}

\textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}. The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}C} and \textbf{D}). Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO (see \textbf{Figure~\ref{fig:kinetics_rec}A}). SSI is also shifted $\sim$ 10 mV to the right (see \textbf{Figure~\ref{fig:kinetics_rec}B}). The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO. Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}E}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}). It is worth mentioning that the proposed method extrapolates $\tau_{i}^{(2)}$ that is greater than 5 s from the 4.5-s-pulse protocol.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kinetics_rec.pdf}
    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, (D) inactivation time constant, (E) I\textsubscript{Kslow2} inactivation time constant, and (F) I\textsubscript{Kss} activation time constant.}
    \label{fig:kinetics_rec}
\end{figure}

We aggregated calibration results into one to analyze the parameters in a collective way. It resulted in a $61\times20$ matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters). Then we applied t-distributed stochastic neighbor (t-SNE) embedding to encapsulate the high-dimensional data into three-dimensional space so that we could visualize each data point. \textbf{Figure~\ref{fig:embd_result}} presents the visualization of the t-SNE embedding of the calibration results. It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups and across the entire dataset.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/embd_result.pdf}
    \caption{t-SNE embedding of calibration parameters.}
    \label{fig:embd_result}
\end{figure}

\section{Conclusions}\label{s:conclusions}
We propose a novel \textit{concurrent data assimilation} framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling. In this study, we propose a novel \textit{concurrent data assimilation} method for modeling K\textsubscript{v} kinetics from multiple I\textsubscript{v} recordings measured from the same cell with different conditions for investigating disease-altered functionalities in the ion channels. This approach provides a complementary tool to physical experiments that are limited in their ability to determine molecular-level channel kinetics rigorously and provide predictive insights. Traditional methods, based on curve-fitting, do not consider interactions of various gating variables across isoforms and dynamics of kinetics according to changes in 
clamp voltages. The proposed method focuses on more comprehensive modeling of underlying channel kinetics of K\textsubscript{v} isoforms than traditional approaches. Experimental results show the strong potential of the proposed framework to discover channel kinetics from various angles. Predictions of kinetics variables to voltage relationships provide information about changes in the voltage dependence of channels. In addition, calibrated parameters of the computer models present a micro-level analysis of channel gating. The embedding method allows visualizing the variability of each cell. The proposed method focuses on the mouse model, but the same framework is applicable to other species easily with modifications to model structures. In summary, the proposed framework paves a new way for investigating the electrical signaling of potassium channels and heart diseases.

\if0\blind{
\section*{Acknowledgements}
This work was supported in part by the National Science Foundation (MCB-1856132 to HY) and (MCB-1856199, IOS-1146882, and IOS-1660926 to EB), an American Heart Association, Greater Southeast Affiliate Grant-In-Aid (14GRNT20450148 to EB), and Postdoctoral Fellowship (15POST25710010 to AE). HK and HY would also like to thank the NSF I/UCRC Center for Health Organization Transformation (CHOT) award NSF IIP-1624727 for the support of their research work.
} \fi

\bibliographystyle{chicago}
\spacingset{1}
\bibliography{IISE-Trans}

\end{document}